Envelope Estimation using Geometric Properties of a Discrete Real Signal

### 2021-05-07 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
formatted according to elsevier template

\documentclass{elsarticle}
% \usepackage[natbib=true]{biblatex}
\usepackage[margin=0.5in]{geometry}
% \usepackage[utf8]{inputenc}
% \usepackage{graphicx}
% \usepackage[backend=biber]{biblatex}
\usepackage{amsmath}
\usepackage{amsfonts} % \mathbb
\usepackage{amssymb}  % \therefore
\usepackage{algorithm}
\usepackage{algorithmic}

@@

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/00ConvexHull.pdf}
  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, and the convex and concave hull at $ \alpha=300 $ defined by it.}
  \label{fig:Hulls}
\end{figure}

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/01AnalyticSignal.pdf}
  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach. In the first half, the sinusoid is free of noise, while in the second half Gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
  \label{fig:AnalyticSignal}
\end{figure}

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/01AnalyticSignalAlto.pdf}
  \caption{Representation of an alto singer sustaining a steady note and a comparison of the envelopes obtained by the Hilbert transform with and without further filtering.}
  \label{fig:AnalyticSignalAlto}
\end{figure}

@@

\begin{subequations}\label{eq:Envelope}
\begin{align}
  \text{Given:}& \\
  \mathbf{w} & \in \mathbb{R}^n, \\
  \mathbf{w} & = (w_0, w_1, \cdots, w_{n-1}) \\
  \text{We define:}& \\
  \mathbf{e}, \mathbf{c} & \in \mathbb{R}^n \\
  \mathbf{e} & = (e_0, e_1, \cdots, e_{n-1}), w_i \ge 0 \ \forall \ i, \ 0 \le i \le n-1 \\  
  \mathbf{c} & = (c_0, c_1, \cdots, c_{n-1}), |c_i| \le 1 \ \forall \ i, \ 0 \le i \le n-1 \\
  \mathbf{w} & = \mathbf{e} \odot \mathbf{c} \ \therefore \ w_i = e_i c_i \ \forall \ i, \ 0 \le i \le n-1
\end{align}
\end{subequations}

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/02Envelope.pdf}
  \caption{Example of a discrete wave \textbf{w} composed by the element-wise multiplication of a previously known envelope \textbf{e} and carrier \textbf{c}. The local extrema of \textbf{w} are highlighted with a circle.}
  \label{fig:Envelope}
\end{figure}

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/03Pulses.pdf}
  \caption{Example of a discrete wave \textbf{w} divided into pulses, of which the extrema are highlighted with a circle. P is the set of the points $ P_j = (i_j, \lvert w_{i_j} \lvert) $ representing the absolute value of those extrema.}
  \label{fig:Pulses}
\end{figure}

@@

\begin{subequations}  \label{eq:Basis}
\begin{align}
P_{\mathbf{b}} &= (x_1, x_2) \\
\beta &= \frac{\left( \frac{i_{m-1} - i_m}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
P_{\mathbf{s}} &= (x_1, \beta x_2) \quad \forall \  P \in \text{P} \\ 
\end{align}
\end{subequations}

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/04Coordinates.pdf}
  \caption{The set of points P, in both the original (top) and standard (bottom) coordinate basis, shown in scale.}
  \label{fig:Coordinates}
\end{figure}

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/04Vectors.pdf}
  \caption{The set of points P and the set V of vectors between two adjacent points of P, in the normalized coordinate system.}
  \label{fig:Vectors}
\end{figure}

@@

\begin{equation} \label{eq:Vectors}
\text{V} = \{ v_0, v_1, \cdots, v_{m-2} \} \quad \text{where} \quad v_k = P_{j+1} - P_j \quad \forall \quad 0 \le k \le m-2
\end{equation}

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/05DiscreteCurvature.pdf}
  \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta_k $ in $ \hat{u}_1 $, $ \theta_k $ being the angle that the vector $ v_k $ makes with the horizontal direction.}
  \label{fig:DiscreteCurvature}
\end{figure}

@@

\begin{equation} \label{eq:Radius}
  r_k = \frac{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}}{ v_{k,y}}
\end{equation}

@@

\begin{equation} \label{eq:Curvature}
\overline{r} = \left( \frac{\sum\limits_{k=0}^{m-2} \left( \frac{ v_{k,y}}{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}} \right)}{m-2} \right)^{-1}
\end{equation}

@@

\begin{subequations} \label{eq:InEnvelope}
\begin{align}
& P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2, \\
& \forall \ P_b, P_c \mid (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2, \\
& P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_b \ne P_c \quad \text{and} \quad x_o, y_o \in \mathbb{R}, \quad 0 \le y_o < +\infty
\end{align}
\end{subequations}

@@

\begin{algorithm}[H]
\caption{Retrieve Envelope} \label{FindEnvelope}
\begin{algorithmic}
\STATE \textbf{Given} $ \text{P} = \{P_0, P_1, \cdots, P_{m-1}\}, r $, \textbf{circle}, a function that returns the center of a circle of a given radius passing through two points and \textbf{distance}, a function that returns the Euclidean distance between two points,
\STATE \textbf{Let} $ \text{E} = \{ P_0 \}, \text{id1} = 0, \text{id2} = 1 $
\STATE \textbf{Define} $ P_c = \text{point}, \text{empty} = \text{boolean} $
\WHILE {$ \text{id2} < m $}
  \STATE $ P_c \leftarrow \textbf{circle}(r, P_\text{id1}, P_\text{id2}) $
  \STATE empty $ \leftarrow $ \TRUE 
  \FOR{$ i = \text{id2} + 1 $ \TO $ m $}
    \IF {$ \textbf{distance}(P_c, P_i) < r $}
      \STATE empty $ \leftarrow $ \FALSE
      \STATE id2 $ \leftarrow \text{id2} + 1 $
      \STATE \textbf{break}
    \ENDIF
  \ENDFOR
  \IF {empty}
    \STATE E $ \leftarrow $ E $ \cup \ \{ P_\text{id2} \} $
    \STATE id1 $ \leftarrow $ id2
    \STATE id2 $ \leftarrow \text{id2} + 1 $
  \ENDIF
\ENDWHILE
\end{algorithmic}
\end{algorithm}

@@

Figs. \ref{fig:spoken_voice}, \ref{fig:singer_voice} and \ref{fig:Bend} present discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier \textbf{c}, obtained after dividing, element-wise, the original signal by its extracted.

From Fig. \ref{fig:spoken_voice} it can be seen that the algorithm performs well under the carrier metric. This is especially remarkable for the case of spoken voice, a notorious complex wave with fast changes in pitch and high noise content.

\begin{figure}[H]
  \centering
    \includegraphics{images/06Graph_spoken_voice.pdf}
  \caption{Signal, envelope and carrier for a record of male voice uttering the word “amazing”.}
  \label{fig:spoken_voice}
\end{figure}

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/06Graph_alto.pdf}
  \caption{Signal, envelope and carrier for a record of an alto singer.}
  \label{fig:singer_voice}
\end{figure}

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/06Graph_bend.pdf}
  \caption{Signal, envelope and carrier for a record of a bend performed on an electric guitar.}
  \label{fig:Bend}
\end{figure}

@@

\begin{subequations} \label{eq:Req1}
\begin{align}
\text{P} &= \{ P_0, P_1, \cdots, P_{m-1} \} \quad \text{where} \quad P_j = (i_j, \lvert w_{i_j} \lvert) \quad \forall \quad 0 \le j \le m-1 \\
\text{E} &\subset \text{P} \implies \max (\text{E}) \le \max (\text{P}) = \max (|w_0|, |w_1|, \cdots, |w_{n-1}|)
\end{align}
\end{subequations}

@@

\begin{subequations}\label{eq:DFT}
\begin{align}
m_\mathbf{w}(f) &= \sum_{k=0}^{n-1} \sqrt{\left( w_k \cos(2 \pi k f / n) \right)^2 + \left( w_k \sin(2 \pi k f / n) \right)^2} \\
m_\mathbf{c}(f) &= \sum_{k=0}^{n-1} \sqrt{\left( \frac{w_k}{e_k} \cos(2 \pi k f / n) \right)^2 + \left( \frac{w_k}{e_k} \sin(2 \pi k f / n) \right)^2} 
\end{align}
\end{subequations}

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/09RefEnvelopeTheory.pdf}
  \caption{Envelope as the boundary of the region filled by a family of curves, for a family of sinusoids with the same frequency and amplitude, modulated by a polynomial.}
  \label{fig:RefEnvelopeTheory}
\end{figure}

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/10RefEnvelope.pdf}
  \caption{Reference envelope obtained by shifting the original signal, for part of a sample of a tom.}
  \label{fig:RefEnvelope}
\end{figure}

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/07Comparison_sinusoid.pdf}
  \caption{Comparison of envelope detection algorithms for a simple sinusoid. The envelope for the Hilbert transform and for the present work are coincident in the image.}
  \label{fig:Comparison_sinusoid}
\end{figure}

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/07Comparison_piano.pdf}
  \caption{Comparison of envelope detection algorithms for the recording of the key 33 of a grand piano}
  \label{fig:Comparison_piano}
\end{figure}

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/07Comparison_soprano.pdf}
  \caption{Comparison of envelope detection algorithms for the recording of the vocalization of a soprano singer}
  \label{fig:Comparison_soprano}
\end{figure}

@@

\begin{table}[H]
\centering
\begin{tabular}{lcccc}
\hline
\multicolumn{5}{c}{\textbf{Average error per frame}} \\
Signal & Present Work & Smoothing & Low pass & Hilbert \\ 
\hline
alto  & 0.001 & 0.337 & 0.342 & 0.163 \\
bend  & 0.000 & 0.029 & 0.030 & 0.008 \\
brass & 0.001 & 0.107 & 0.107 & 0.038 \\
nonperiodic & 0.000 & 0.211 & 0.158 & 0.000 \\
piano & 0.003 & 0.016 & 0.016 & 0.007 \\
sinusoid & 0.000 & 0.132 & 0.131 & 0.000 \\
soprano & 0.000 & 0.155 & 0.155 & 0.044 \\
spoken\_voice & 0.006 & 0.109 & 0.110 & 0.045 \\
tom   & 0.001 & 0.011 & 0.015 & 0.005 \\
\hline
Average & 0.001 & 0.123 & 0.118 & 0.035 \\
frac. of max. & 0.010 & 1.000 & 0.961 & 0.281 \\
\end{tabular}
\caption{Comparison of the average errors of the algorithm presented in this work with the most common methods of digital envelope identification. The discrete waves used were normalized between -1 and 1. The average is between the absolute value of those waves and the envelope divided by two.}
\label{table:Error_comparison}
\end{table}

@@

\begin{table}[H]
\centering
\begin{tabular}{lcccc}
\hline
\multicolumn{5}{c}{\textbf{Time in seconds}} \\
Signal  & Present Work & Smoothing & Low pass & Hilbert \\ 
\hline
alto  & 0.010 & 0.096 & 0.007 & 0.006 \\
bend  & 0.009 & 0.146 & 0.005 & 0.067 \\
brass & 0.015 & 0.253 & 0.005 & 0.035 \\
nonperiodic & 0.002 & 0.064 & 0.003 & 0.007 \\
piano & 0.022 & 0.316 & 0.009 & 0.030 \\
sinusoid & 0.003 & 0.069 & 0.003 & 0.006 \\
soprano & 0.111 & 0.574 & 0.013 & 0.293 \\
spoken\_voice & 0.003 & 0.062 & 0.003 & 0.017 \\
tom   & 0.002 & 0.070 & 0.001 & 0.004 \\
\hline
Average & 0.020 & 0.183 & 0.005 & 0.052 \\
frac. of max. & 0.108 & 1.000 & 0.030 & 0.283 \\
\end{tabular}
\caption{Comparison of the processing time of the algorithm presented in this work with the most common methods of digital envelope detection.}
\label{table:Time_comparison}
\end{table}

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/08FullFrontiers.pdf}
  \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented. For each wave, the region highlighted in black is shown in detail besides the whole wave. For each wave, the horizontal axis is the sample number $ i $, while the vertical axis is the normalized amplitude. All 6 waves were recorded at 44100 fps.}
  \label{fig:FullFrontiers}
\end{figure}

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/11Fourier.pdf}
  \caption{Fourier power spectrum for the wave and carrier shown in \ref{fig:Envelope}.}
  \label{fig:Fourier}
\end{figure}

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/12Fourier.pdf}
  \caption{Fourier power spectrum for the wave and carrier of the sound of a guitar bend shown in Fig. \ref{fig:Bend}.}
  \label{fig:FourierBend}
\end{figure}

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/09PCs.pdf}
  \caption{Section of the digital wave illustrated in \ref{fig:singer_voice}, with vertical lines at the indices of the samples that belong to the positive frontier.}
  \label{fig:PCs}
\end{figure}

@@

\begin{figure}[H]
  \centering
    \includegraphics{images/10AvgPc.pdf}
  \caption{Pseudo-cycles of the wave illustrated in \ref{fig:singer_voice}, segmented and superposed, and their average. The average is also shown shifted by one average period, to illustrate the average continuity.}
  \label{fig:AvgPc}
\end{figure}

>>>>

\documentclass[preprint]{elsarticle}
% \usepackage[natbib=true]{biblatex}
\usepackage[margin=0.5in]{geometry}
% \usepackage[utf8]{inputenc}
% \usepackage{graphicx}
% \usepackage[backend=biber]{biblatex}
\usepackage{amsmath}
\usepackage{amsfonts} % \mathbb
\usepackage{amssymb}  % \therefore
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{lineno,hyperref}
% \modulolinenumbers[5]

@@

\linenumbers

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/00ConvexHull.pdf}
  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, and the convex and concave hull at $ \alpha=300 $ defined by it.}
  \label{fig:Hulls}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/01AnalyticSignal.pdf}
  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach. In the first half, the sinusoid is free of noise, while in the second half Gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
  \label{fig:AnalyticSignal}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/01AnalyticSignalAlto.pdf}
  \caption{Representation of an alto singer sustaining a steady note and a comparison of the envelopes obtained by the Hilbert transform with and without further filtering.}
  \label{fig:AnalyticSignalAlto}
\end{figure}

@@

\begin{align} \label{eq:Envelope}
\begin{split}
  \text{Given:}& \\
  \mathbf{w} & \in \mathbb{R}^n, \\
  \mathbf{w} & = (w_0, w_1, \cdots, w_{n-1}) \\
  \text{We define:}& \\
  \mathbf{e}, \mathbf{c} & \in \mathbb{R}^n \\
  \mathbf{e} & = (e_0, e_1, \cdots, e_{n-1}), w_i \ge 0 \ \forall \ i, \ 0 \le i \le n-1 \\  
  \mathbf{c} & = (c_0, c_1, \cdots, c_{n-1}), |c_i| \le 1 \ \forall \ i, \ 0 \le i \le n-1 \\
  \mathbf{w} & = \mathbf{e} \odot \mathbf{c} \ \therefore \ w_i = e_i c_i \ \forall \ i, \ 0 \le i \le n-1
\end{split}
\end{align}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/02Envelope.pdf}
  \caption{Example of a discrete wave \textbf{w} composed by the element-wise multiplication of a previously known envelope \textbf{e} and carrier \textbf{c}. The local extrema of \textbf{w} are highlighted with a circle.}
  \label{fig:Envelope}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/03Pulses.pdf}
  \caption{Example of a discrete wave \textbf{w} divided into pulses, of which the extrema are highlighted with a circle. P is the set of the points $ P_j = (i_j, \lvert w_{i_j} \lvert) $ representing the absolute value of those extrema.}
  \label{fig:Pulses}
\end{figure}

@@

\begin{align}\label{eq:Basis}
\begin{split}
P_{\mathbf{b}} &= (x_1, x_2) \\
\beta &= \frac{\left( \frac{i_{m-1} - i_m}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
P_{\mathbf{s}} &= (x_1, \beta x_2) \quad \forall \  P \in \text{P} \\ 
\end{split}
\end{align}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/04Coordinates.pdf}
  \caption{The set of points P, in both the original (top) and standard (bottom) coordinate basis, shown in scale.}
  \label{fig:Coordinates}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/04Vectors.pdf}
  \caption{The set of points P and the set V of vectors between two adjacent points of P, in the normalized coordinate system.}
  \label{fig:Vectors}
\end{figure}

@@

\begin{align} \label{eq:Vectors}
\text{V} = \{ v_0, v_1, \cdots, v_{m-2} \} \quad \text{where} \quad v_k = P_{j+1} - P_j \quad \forall \quad 0 \le k \le m-2
\end{align}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/05DiscreteCurvature.pdf}
  \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta_k $ in $ \hat{u}_1 $, $ \theta_k $ being the angle that the vector $ v_k $ makes with the horizontal direction.}
  \label{fig:DiscreteCurvature}
\end{figure}

@@

\begin{align} \label{eq:Radius}
  r_k = \frac{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}}{ v_{k,y}}
\end{align}

@@

\begin{align} \label{eq:Curvature}
\overline{r} = \left( \frac{\sum\limits_{k=0}^{m-2} \left( \frac{ v_{k,y}}{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}} \right)}{m-2} \right)^{-1}
\end{align}

@@

\begin{align} \label{eq:InEnvelope}
\begin{split}
& P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2, \\
& \forall \ P_b, P_c \mid (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2, \\
& P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_b \ne P_c \quad \text{and} \quad x_o, y_o \in \mathbb{R}, \quad 0 \le y_o < +\infty
\end{split}
\end{align}

@@

\begin{algorithm}[ht!]
\caption{Retrieve Envelope} \label{FindEnvelope}
\begin{algorithmic}
\STATE \textbf{Given} $ \text{P} = \{P_0, P_1, \cdots, P_{m-1}\}, r $, \textbf{circle}, a function that returns the center of a circle of a given radius passing through two points and \textbf{distance}, a function that returns the Euclidean distance between two points,
\STATE \textbf{Let} $ \text{E} = \{ P_0 \}, \text{id1} = 0, \text{id2} = 1 $
\STATE \textbf{Define} $ P_c = \text{point}, \text{empty} = \text{boolean} $
\WHILE {$ \text{id2} < m $}
  \STATE $ P_c \leftarrow \textbf{circle}(r, P_\text{id1}, P_\text{id2}) $
  \STATE empty $ \leftarrow $ \TRUE 
  \FOR{$ i = \text{id2} + 1 $ \TO $ m $}
    \IF {$ \textbf{distance}(P_c, P_i) < r $}
      \STATE empty $ \leftarrow $ \FALSE
      \STATE id2 $ \leftarrow \text{id2} + 1 $
      \STATE \textbf{break}
    \ENDIF
  \ENDFOR
  \IF {empty}
    \STATE E $ \leftarrow $ E $ \cup \ \{ P_\text{id2} \} $
    \STATE id1 $ \leftarrow $ id2
    \STATE id2 $ \leftarrow \text{id2} + 1 $
  \ENDIF
\ENDWHILE
\end{algorithmic}
\end{algorithm}

@@

Figs. \ref{fig:spokenvoice}, \ref{fig:singer_voice} and \ref{fig:Bend} present discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier \textbf{c}, obtained after dividing, element-wise, the original signal by its extracted.

From Fig. \ref{fig:spokenvoice} it can be seen that the algorithm performs well under the carrier metric. This is especially remarkable for the case of spoken voice, a notorious complex wave with fast changes in pitch and high noise content.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/06Graph_spoken_voice.pdf}
  \caption{Signal, envelope and carrier for a record of male voice uttering the word “amazing”.}
  \label{fig:spokenvoice}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/06Graph_alto.pdf}
  \caption{Signal, envelope and carrier for a record of an alto singer.}
  \label{fig:singer_voice}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/06Graph_bend.pdf}
  \caption{Signal, envelope and carrier for a record of a bend performed on an electric guitar.}
  \label{fig:Bend}
\end{figure}

@@

\begin{align} \label{eq:Req1}
\begin{split}
\text{P} &= \{ P_0, P_1, \cdots, P_{m-1} \} \quad \text{where} \quad P_j = (i_j, \lvert w_{i_j} \lvert) \quad \forall \quad 0 \le j \le m-1 \\
\text{E} &\subset \text{P} \implies \max (\text{E}) \le \max (\text{P}) = \max (|w_0|, |w_1|, \cdots, |w_{n-1}|)
\end{split}
\end{align}

@@

\begin{align} \label{eq:DFT}
\begin{split}
m_\mathbf{w}(f) &= \sum_{k=0}^{n-1} \sqrt{\left( w_k \cos(2 \pi k f / n) \right)^2 + \left( w_k \sin(2 \pi k f / n) \right)^2} \\
m_\mathbf{c}(f) &= \sum_{k=0}^{n-1} \sqrt{\left( \frac{w_k}{e_k} \cos(2 \pi k f / n) \right)^2 + \left( \frac{w_k}{e_k} \sin(2 \pi k f / n) \right)^2} 
\end{split}
\end{align}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/09RefEnvelopeTheory.pdf}
  \caption{Envelope as the boundary of the region filled by a family of curves, for a family of sinusoids with the same frequency and amplitude, modulated by a polynomial.}
  \label{fig:RefEnvelopeTheory}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/10RefEnvelope.pdf}
  \caption{Reference envelope obtained by shifting the original signal, for part of a sample of a tom.}
  \label{fig:RefEnvelope}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/07Comparison_sinusoid.pdf}
  \caption{Comparison of envelope detection algorithms for a simple sinusoid. The envelope for the Hilbert transform and for the present work are coincident in the image.}
  \label{fig:Comparison_sinusoid}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/07Comparison_piano.pdf}
  \caption{Comparison of envelope detection algorithms for the recording of the key 33 of a grand piano}
  \label{fig:Comparison_piano}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/07Comparison_soprano.pdf}
  \caption{Comparison of envelope detection algorithms for the recording of the vocalization of a soprano singer}
  \label{fig:Comparison_soprano}
\end{figure}

@@

\begin{table}[ht!]
\centering
\begin{tabular}{lcccc}
\hline
\multicolumn{5}{c}{\textbf{Average error per frame}} \\
Signal & Present Work & Smoothing & Low pass & Hilbert \\ 
\hline
alto  & 0.001 & 0.337 & 0.342 & 0.163 \\
bend  & 0.000 & 0.029 & 0.030 & 0.008 \\
brass & 0.001 & 0.107 & 0.107 & 0.038 \\
nonperiodic & 0.000 & 0.211 & 0.158 & 0.000 \\
piano & 0.003 & 0.016 & 0.016 & 0.007 \\
sinusoid & 0.000 & 0.132 & 0.131 & 0.000 \\
soprano & 0.000 & 0.155 & 0.155 & 0.044 \\
spoken\_voice & 0.006 & 0.109 & 0.110 & 0.045 \\
tom   & 0.001 & 0.011 & 0.015 & 0.005 \\
\hline
Average & 0.001 & 0.123 & 0.118 & 0.035 \\
frac. of max. & 0.010 & 1.000 & 0.961 & 0.281 \\
\end{tabular}
\caption{Comparison of the average errors of the algorithm presented in this work with the most common methods of digital envelope identification. The discrete waves used were normalized between -1 and 1. The average is between the absolute value of those waves and the envelope divided by two.}
\label{table:Error_comparison}
\end{table}

@@

\begin{table}[ht!]
\centering
\begin{tabular}{lcccc}
\hline
\multicolumn{5}{c}{\textbf{Time in seconds}} \\
Signal  & Present Work & Smoothing & Low pass & Hilbert \\ 
\hline
alto  & 0.010 & 0.096 & 0.007 & 0.006 \\
bend  & 0.009 & 0.146 & 0.005 & 0.067 \\
brass & 0.015 & 0.253 & 0.005 & 0.035 \\
nonperiodic & 0.002 & 0.064 & 0.003 & 0.007 \\
piano & 0.022 & 0.316 & 0.009 & 0.030 \\
sinusoid & 0.003 & 0.069 & 0.003 & 0.006 \\
soprano & 0.111 & 0.574 & 0.013 & 0.293 \\
spoken\_voice & 0.003 & 0.062 & 0.003 & 0.017 \\
tom   & 0.002 & 0.070 & 0.001 & 0.004 \\
\hline
Average & 0.020 & 0.183 & 0.005 & 0.052 \\
frac. of max. & 0.108 & 1.000 & 0.030 & 0.283 \\
\end{tabular}
\caption{Comparison of the processing time of the algorithm presented in this work with the most common methods of digital envelope detection.}
\label{table:Time_comparison}
\end{table}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/08FullFrontiers.pdf}
  \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented. For each wave, the region highlighted in black is shown in detail besides the whole wave. For each wave, the horizontal axis is the sample number $ i $, while the vertical axis is the normalized amplitude. All 6 waves were recorded at 44100 fps.}
  \label{fig:FullFrontiers}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/11Fourier.pdf}
  \caption{Fourier power spectrum for the wave and carrier shown in \ref{fig:Envelope}.}
  \label{fig:Fourier}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/12Fourier.pdf}
  \caption{Fourier power spectrum for the wave and carrier of the sound of a guitar bend shown in Fig. \ref{fig:Bend}.}
  \label{fig:FourierBend}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/09PCs.pdf}
  \caption{Section of the digital wave illustrated in \ref{fig:singer_voice}, with vertical lines at the indices of the samples that belong to the positive frontier.}
  \label{fig:PCs}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/10AvgPc.pdf}
  \caption{Pseudo-cycles of the wave illustrated in \ref{fig:singer_voice}, segmented and superposed, and their average. The average is also shown shifted by one average period, to illustrate the average continuity.}
  \label{fig:AvgPc}
\end{figure}
